---
title: "YK_Assignment5"
author: "Yinan Kang"
date: "3/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 4.9 - PLASTIC HARDNESS 
## (a)
```{r}
# Import plastic hardness data 

colnames <- c("y","x") 
df <- read.table(url("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR22.txt"), header=FALSE, sep="", col.names=colnames)
n <- nrow(df) 

# (a)
# Bonferroni confidence intervals for X = 20, 30, 40 hours
fitreg <- lm(y~x, df)
Xh <- c(20,30,40)
pred <- predict.lm(fitreg,data.frame(x = Xh), se.fit = T, level = 0.9)
B = rep(qt(1-0.1/(2*length(Xh)),n-2),length(Xh))
a.results <- cbind(pred$fit-B*pred$se.fit, pred$fit+B*pred$se.fit)
colnames(a.results) <- c("lwr","upr")

print(a.results) 
```
What is the meaning of the family confidence coefficient here?  

Response: Family confidence coefficient at 90% confidence means, when sampled repeatedly, 90% of the time the entire family of estimates would be valid.  

Here specifically, if we sampled the below Xh values repeatedly, 90% of the time all three Yh values would exist in their respective ranges:   
Xh = 20 (hours), the Yh (hardness) value between [`r a.results[1,1]`, `r a.results[1,2]`]  
Xh = 30, the Yh value between [`r a.results[2,1]`, `r a.results[2,2]`]  
Xh = 40, the Yh value between [`r a.results[3,1]`, `r a.results[3,2]`]  


## (b) 
```{r}
# Checking Working-Hotelling interval - if it produces a tighter prediction interval, it'd be considered more efficient 

W <- rep(sqrt(2*qf(0.9,2,n-2)),length(Xh))
b.results <- cbind(pred$fit-W*pred$se.fit, pred$fit + W*pred$se.fit)
colnames(b.results) <- c("lwr","upr")
print(b.results)
```
Is the Bonferroni procedure the most efficient one to be employed here? 

Response: No, the Working-Hotelling procedure produced tighter intervals among the Yh predictions shown thus far. Therefore, the Working-Hotelling procedure is the more efficient one here. 

## (c) 
```{r}
# Predict Yh for Xh = 30,40 using Working-Hotelling at 90% confidence

Xh <- c(30,40)
pred <- predict.lm(fitreg,data.frame(x = Xh), se.fit = T, level = 0.9)
W <- rep(sqrt(2*qf(0.9,2,n-2)),length(Xh))
c.results <- cbind(pred$fit-W*pred$se.fit, pred$fit + W*pred$se.fit)
colnames(c.results) <- c("lwr","upr")
print(c.results)
```

Predictions of Yh for Xh = 30,40 with 90% Confidence:  

When Xh = 30, Yh is between `r c.results[1,1]` and `r c.results[1,2]`  
When Xh = 40, Yh is between `r c.results[2,1]` and `r c.results[2,2]`  

# Problem 4.19 - GPA 
## (a)
```{r}
# Import Data
rm(list=ls())
colnames <- c("y","x") 
df <- read.table(url("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR19.txt"), header=FALSE, sep="", col.names=colnames)
n <- nrow(df) 

# Create linear model
model <- lm(y~x, df)

# Collecting figures needed for s2{predX}
mse <- mean(model$residuals^2)
b0 <- 2.11405 # From model
b1 <- 0.03883 # From model
xbar <- mean(df$x)
sxx <- 0
for (i in 1:nrow(df)){
  sxx.temp <- (df$x[i] - xbar)^2
  sxx <- sxx+sxx.temp
}

# Calculating Xh.new
Yh.new <- 3.4
Xh.new <- (Yh.new-b0)/b1

# Calculating s{predx} for Xh = 3.4 
s2.predx <- (mse/(b1^2)) * (1+(1/n) + ((Xh.new-xbar)^2)/sxx)
s.predx <- sqrt(s2.predx) 

# Calculating Interval at 90%
t <- qt(1-0.1/2,n-2)

Xh.new.low <- Xh.new-t*s.predx
Xh.new.hi <- Xh.new + t*s.predx

print(Xh.new.low)
print(Xh.new.hi)
```

Results: For a student with GPA = 3.4, we are 90% confident that their ACT score would've been between `r Xh.new.low` and `r Xh.new.hi`. 

# (b)
```{r}
# Checking criterion 4.33
crit <- ( (t^2) * mse ) / ((b1^2) * sxx) 
print(crit) 

```
The guideline given for criterion (4.33) is it should be small, < 0.1. Here, the value is 0.29, therefore the criterion for appropriateness is NOT MET. 

# Problem 4.26 - CDI 
## (a) Obtain joint Bonferroni confidence intervals at 95%

```{r, message = FALSE,warning= FALSE}
# Importing Data
rm(list=ls())
cdi.df <- read.csv("CDI.csv") 
n <- nrow(cdi.df)
attach(cdi.df) 

# Generating linear regression model
model <- lm(Number.Active.Physicians ~ Total.Population)
b0 <- -1.106e+02   # from model
b1 <- 2.795e-03   # from model

# Calculating B 
B <- qt(1-.05/4,n-2)

# Collecting values for s.b0 and s.b1 (to calculate confidence intervals)
mse <- mean(model$residuals^2)
xbar <- mean(Total.Population)
sxx <- 0
for (i in 1:nrow(cdi.df)) {
  sxx.temp <- (Total.Population[i] - xbar)^2
  sxx <- sxx + sxx.temp
}
s2.b0 <- mse * ((1/n) + (xbar^2)/sxx)
s.b0 <- sqrt(s2.b0)
s2.b1 <- mse/sxx
s.b1 <- sqrt(s2.b1)

# Calculating confidence intervals 

b0.low <- b0 - B*s.b0
b0.hi <- b0 + B*s.b0
b1.low <- b1 - B*s.b1
b1.hi <- b1 + B*s.b1
```
Bonferroni Joint Confidence Intervals:  
b0: [$`r b0.low`$, $`r b0.hi`$]  
b1: [$`r b1.low`$, $`r b1.hi`$]  

## (b) 
Do the joint confidence intervals from (a) support the view that B0 should be -100 and B1 should be .0028? 

Response: YES, according to the confidence intervals for b0 and b1 from (a), the researcher's views on beta0 and beta1 can be supported. 

## (c)
Will the Working-Hotelling or Bonferroni procedure be more efficient?

Reponse: Based on the reasoning provided in the textbook, for smaller families of prediction with fewer statements, the Bonferroni will give the more efficient and tighter interval. This is the case here (3 prediction intervals), so BONFERRONI would likely be more efficient. 

## (d)
```{r}
# Generating Bonferroni 90% confidence intervals
Xh <- c(500*1000, 1000*1000, 5000*1000)
pred <- predict.lm(model, data.frame(Total.Population=Xh),se.fit=T,level=0.9)

B <- rep(qt(1-0.1/(2*length(Xh)),n-2),length(Xh))

d.results <- cbind(pred$fit-B*pred$se.fit, pred$fit+B*pred$se.fit)
colnames(d.results) <- c("lwr","upr")
print(d.results)

```


Interpretation: 

When repeatedly sampled, 90% of the time areas with the following 'populations' will have 'number of physicians' in the given ranges:   
For populations of 500 thousand, the mean number of physicians is in range: [$`r d.results[1,1]`$, $`r d.results[1,2]`$]   
For populations of 1000 thousand, the mean number of physicians is in range: [$`r d.results[2,1]`$, $`r d.results[2,2]`$]   
For populations of 5000 thousand, the mean number of physicians is in range: [$`r d.results[3,1]`$, $`r d.results[3,2]`$]   


# Problem 4.27 - SENIC 

## (a)
```{r message=FALSE, warning=FALSE}
rm(list=ls())
# Import Data
senic <- read.table("/cloud/project/APC1.DAT", quote="\"", comment.char="")
names(senic) <- c("ID", "length.of.stay", "age" , "infection.risk", "routine.culturing.ratio", "routine.chest.xray.ratio", "number.of.beds","medical.school.affiliation", "region","average.daily.census","number.of.nurses","available.facilities.services")
n <- nrow(senic) 

attach(senic) 

# Generating model
model <- lm(length.of.stay ~ infection.risk)

# Bonferroni join confidence intervals for B0 and B1, 90% family confidence 
bonf <- confint(model,level=1-0.1/2)
print(bonf) 
```

Joint confidence interval, 90% Bonferroni:  
B0: [`r bonf[1,1]`, `r bonf[1,2]`]  
B1: [`r bonf[2,1]`, `r bonf[2,2]`]  

## (b) 
Does joint confidence interval in (a) support researcher's belief that B0 = 7, B1 = 1 ? 

Response: No, the joint confidence interval in (a) DOES NOT support the researcher's belief, as B1 = 1 is not in the joint confidence interval range for B1. 

## (c)
Will the Working-Hotelling or Bonferroni procedure be more efficient for creating CIs for new Xh values? 

Response: Not sure here, as Xh has size 4, which is not tiny, thus not guaranteeing Bonferroni would be more efficient. Thus, I'll try both in (d) and see for myself. 

## (d) 
```{r}
Xh <- c(2,3,4,5)
pred <- predict.lm(model,data.frame(infection.risk=Xh),se.fit=T,level=.95)    # 95% confidence interval

# Creating Bonferroni 
B <- rep(qt(1-.05/(2*length(Xh)),n-2),length(Xh))
b.results <- cbind(pred$fit-B*pred$se.fit, pred$fit+B*pred$se.fit)

# Creating Working-Hotelling 
W <- rep(sqrt(2*qf(0.95,2,n-2)),length(Xh))
w.results <- cbind(pred$fit-W*pred$se.fit, pred$fit+W*pred$se.fit)

# Compare Bonferroni vs. Working-Hotelling 
print(b.results)
print(w.results) 

```

Conclusion: WORKING-HOTELLING was the more efficient choice here. This makes sense, as the family size = 4, which is large enough for the Working-Hotelling to be tighter than the Bonferroni. 

Interpretation: When repeatedly sampled, 95% of the time the following family of estimates will be valid:  

Infection Risk = 2, Avg. Length of Hospital Stay = [`r w.results[1,1]`, `r w.results[1,2]`]  
Infection Risk = 3, Avg. Length of Hospital Stay = [`r w.results[2,1]`, `r w.results[2,2]`]  
Infection Risk = 4, Avg. Length of Hospital Stay = [`r w.results[3,1]`, `r w.results[3,2]`]  
Infection Risk = 5, Avg. Length of Hospital Stay = [`r w.results[4,1]`, `r w.results[4,2]`]  



