---
title: "Assignment2"
author: "Yinan Kang"
date: "2/10/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Problem 1.22
## (a)
```{r}
batches.df <- data.frame(X = c(16,16,16,16,24,24,24,24,32,32,32,32,40,40,40,40), 
                Y = c(199,205,196,200,218,220,215,223,237,234,235,230,250,248,253,246))
head(batches.df,3)

set.seed(258) 

batches.lm <- lm(batches.df$Y ~ batches.df$X)
batches.lm
```
The estimated linear regression function is Y = 2.034*X + 168.6.

```{r}
plot(batches.df)
abline(168.6, 2.034)
```

Yes, the regression function looks like a good fit here as the values tend to be clustered near the line. 

## (b)
```{r}
#Using X = 40:
Y.40 <- 2.034*40 + 168.6 
print(Y.40)
```
Point estimate at X = 40 is 249.96

## (c)
Point estimate of change in mean hardness when X increases by 1 is equal to the regression coefficient, 2.034 Brinell units 

# Problem 2.16

```{r}
mean((batches.lm$residuals)^2)
mean(batches.df$X)
sum.xtot <- 0
for (i in 1:nrow(batches.df)) {
  sum.xind <- (batches.df[i,1] - mean(batches.df$X))^2
  sum.xtot <- sum.xtot + sum.xind
  
}
```
MSE of linear regression in Problem 1.22 = 9.15  
X-bar = 28 hr  
sum((Xi - X-bar)^2) = 1280  