---
title: "YK_Final_P5"
author: "Yinan Kang"
date: "5/14/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
df <- read.csv("/cloud/project/Question 5.csv")

require(ResourceSelection)
```


# (a) Fit model with first orders plus all interactions
```{r} 
model <- glm(Y~X1+X2+X3+X4+X1*X2+X1*X3+X1*X4+X2*X3+X2*X4+X3*X4, family="binomial",data=df)
summary(model)
```


# (b) Likelihood Ratio Test for dropping Interaction Terms

Full Model: Pi = [1+exp(-X'BetaFull)] ^-1   
where X*BetaFull = Beta0 + Beta1*X1 + .... Beta10*(X3*X4)   

Reduced Model: Pi = [1+exp(-X'BetaReduced)] ^-1  
where X*BetaFull = Beta0 + Beta1*X1 + .... Beta4*X4    


Hypotheses:  
H0: Beta 5 = Beta 6 .... = Beta 10 = 0   (All Beta's with Interaction terms = 0)  
Ha: Not All Betas in H0 = 0   

Decision Rule:  
If G2 <= Chi-Sq(1-alpha, p-q), conclude H0   
If G2 > Chi-Sq(1-alpha, p-1), conclude Ha   

Assume alpha = 0.05  


```{r}
model.r <- glm(Y~X1+X2+X3+X4, family = "binomial", data=df)
anova(model, model.r, test="Chisq")
```


**Decision**:  

We use the 'anova' function to compare our full and reduced models. We see that the p-value associated with the Chi-Sq distribution is > 0.05.   
This interpretation is different than the formal one stated in the Decision Rule above. This p-value lets us know that there is no statistical significance in adding the interactions, so they can be DROPPED.   


## (c) Backward elimination
```{r}
step(model, scope = formula(model), direction="backward")
```


**Analysis**: Using Backward step, we see that X1, X2, X3 are retained.   


## Re-fitting model for rest of problem  !!
```{r}
model <- glm(Y~X1+X2+X3, family="binomial", data=df)

```

## (d) Hoslem-Lameshow Goodness-of-Fit

Hypotheses:  

H0: Logistic response is appropriate  
Ha: Logistic Response is not appropriate  

Alternatives: 

If X.test <= Chi-Sq(0.95,c-2), conclude H0  
If X.test > Chi-Sq(0.95,c-2), conclude Ha  


```{r}
hoslem.test(model$y,fitted(model),g=5)
```


**Decision**: As X.test < Chi-Sq, we conclude H0, that a Logistic response is appropriate.   


## (e) Prediction

```{r}
pred.data <- data.frame(X1=c(33,6), X2 =c(1,1), X3 = c(1,1), X4 = c(0,0))

pred <- predict(model, newdata=pred.data)
print(pred)

# Ran out of time to do joint confidence interval
```